### 2014-02-28 paper/retrofit.tex
Minor changes to retrofit.

\and
\inferrule[T-deref]
{ }
{| conf tS (!ta) -> conf tS (ap ts ta) |}

@@

Information flow control seeks to enforce a security policy known as
\emph{noninterference}: public results must not depend on secret inputs.
Modern, dynamic IFC systems usually achieve this by associating a \emph{label}
with data, which encodes a policy of whether or not data tagged with one
label can flow to |canFlowTo| another label.  Since our embedding is agnostic
to the choice of labeling scheme, we simply represent labels with the metavariable
$l$, but do not discuss them in more detail.

In Figure~\ref{fig:ifc}, we give the syntax and \emph{single-threaded}
evaluation rules for a minimal information-flow control language.  The
evaluation context $\mathcal{E}_{\ifc{\Sigma}}$ represents the context
of the current thread which is running, including on the right-hand side
the current label $l$ (which protects the program counter of the thread)
and the thread ID $i$.

@@

\begin{itemize}
    \item First-class labels, with label values $l$ as well as operations for computing on
labels (|canFlowTo|, |lub| and |glb|),
    \item Operations for inspecting (\textbf{getLabel}) and modifying (\textbf{setLabel}) the current label of the thread (a thread can only increase its label, and not decrease it),
    \item Operations for non-blocking inter-thread communication (\textbf{send} and \textbf{recv}),
        which interact with the global store of per-thread message queues |iS|, and
    \item A forking operation, which will be described in more detail in the next section.
\end{itemize}

The form of \textbf{send}/\textbf{recv} are inspired by methods of inter-worker communication
in web browser APIs \Red{cite}, however in Section~\ref{sec:extensions} we
discuss other possible choices for information flow control aware communication
primitives.  Their rules are non-standard and worth commenting on: when a thread places
a message on the message queue of the thread it is sending to, it also adds the current
label when it sent the message.  When the recipient reads out messages from its queue,
it filters out (notated using |filter il Q|) messages whose labels do not
flow to |il|. Specifically, if |Q| is the empty list |nil|, the
function is simply the identity function, i.e.,
|filter il nil = nil|; otherwise:
\[
|filter il ((il', id, ie) , Q )| = \left\{
\begin{array}{l l}
|(il', id, ie) , filter il Q| & \quad \text{if |il' canFlowTo il|}\\
|filter il Q| & \quad \text{otherwise}
\end{array} \right.
\]

@@

\inferrule[I-labelOp]
{ |denot (il1 lop il2) = iv|}
{ |niconf iS id il (il1 lop il2) -> niconf iS id il iv| }
\end{mathpar}
\caption{IFC language, single-thread operations}
\label{fig:ifc}
\end{figure}

@@

\Red{TODO[ds]:
  - move the async recv to text and out of the figure
  - clarify discussion on exceptions
  - fix the toLabled
  - add threadId to ruls
}
Finally, in Figure~\ref{fig:embedding},
we provide all of the rules responsible for actually carrying out the embedding.
The most important feature of this embedding is that every thread maintains its own
copy of the target language global state and evaluation context, thus
enforcing isolation between various threads.  In more detail:

@@

\begin{itemize}
    \item We extend the values, expressions and evaluation contexts of both languages to allow for embeddings terms in one language to be embedded in the other.
    \item We reinterpret |CE| to be evaluation contexts on threadpools, providing definitions for $\mathcal{E}_{\tar{\Sigma}}$ and $\mathcal{E}_{\ifc{\Sigma}}$.  These rules only operate on the first thread in the threadpool, which by convention, is the only thread executing.
    \item We reinterpret |->|, an operation on a single thread, in terms of |.->|, operation on thread pools.  The correspondence is simple: a thread executes a step and then is rescheduled in the thread pool according to schedule policy $\alpha$.  Not all schedulers preserve non-interference, so we provide two specific schedules in Figure~\ref{fig:scheduler}.
    \item Finally, we define some rules for scheduling, handling forking threads (which interact with the state of the target language) and intermediating between the borders of the two languages.
\end{itemize}

The \textsc{I-fork} rule is of particular interest, as it is intended to allow users
to create new threads.  When a new thread is created, there is a question of what the
global state of the new thread should be; this is defined in terms of a function $\kappa$.
Conservatively, $\kappa$ may be simply thought of as the identity function, in which
case the semantics of fork are such that the state of the target language is \emph{cloned}
when forking occurs.  However, this is not necessary: it is also valid for $\kappa$
to remove entries from the state.  In Section~\ref{sec:concrete}, we give a more detailed
discussion of the implications of the choice of $\kappa$.

@@

\paragraph{Sending mutable references}  As each thread maintains its own
global state for the target language, one might reasonably wonder what would
happen if you attempted to send a reference to another thread.

@@

It would be a violation of information flow if such an operation actually
resulted in references being shared between two threads.  In fact, the address
will be received by the forked thread, but it will be a dangling pointer; the
thread will become stuck and removed from execution by \textsc{I-noStep}.  (Readers
who are curious whether or not a static discipline could be enforced to
prevent these dangling pointers should check Section~\ref{sec:extensions}.)

@@

When $\kappa$ is the identity, the forked thread successfully dereferences
the contents of the reference to be true twice (recall that references
are cloned into the new global state).  When $\kappa$ is the constant function
to the empty store, $x$ becomes a dangling pionter.

@@

At first glance, it does not appear that our system has any such operator;
there are only threads, and the correct behavior for a thread throwing
an exception is self-evident: it should simply be removed from the scheduler.
However, under a \emph{sequential scheduler}, the behavior of \textbf{fork}
can be used to model the behavior of \textbf{toLabeled}:

\begin{align*}
    & |toLabeled l te @= | \\
    & |tab let pid = threadId; | \\
    & |tab let i = TI (fork (IT (TI (setLabel l); TI (send pid l (IT te)))))| \\
    & |tab in (\ _ . TI (setLabel l); TI ( x, _ <= recv . x ))| \\
\end{align*}

When the sequential scheduler is used, the forked thread runs to
completion before returning to the thread that forked it.  Thus,
a threadpool should rather be interpreted as a number of information-flow
aware stack frames.  To read the result of a |toLabeled| operation
(which is communicated via message passing), the parent thread must raise
its label; otherwise the label will get dropped, as per the \textsc{I-recv} rule.\footnote{A keen reader may note that with a single, per-thread
channel, we might mix up our messages.  Indeed, a proper development requires
the ability to allocate fresh channels; for purposes of simplicity, this
has been omitted.}

>>>>

\and
\inferrule[T-deref]
{ }
{| conf tS (!ta) -> conf tS (ap tS ta) |}

@@

Information flow control seeks to enforce a security policy known as
\emph{non-interference}: public results must not depend on secret inputs.
Modern, dynamic IFC systems usually achieve this by associating a \emph{label}
with data, which encodes a policy of whether or not data tagged with one
label can flow to another label (written as |canFlowTo|).
Since our embedding is agnostic
to the choice of labeling scheme, we simply represent labels with the metavariable
$l$, but do not discuss them in more detail.

In Figure~\ref{fig:ifc}, we give the syntax and \emph{single-threaded}
evaluation rules for a minimal information-flow control language.  The
evaluation context $\mathcal{E}_{\ifc{\Sigma}}$ represents the context
of the current task which is running, including on the right-hand side
the current label $l$ (which protects the program counter of the task)
and the task ID $i$.

@@

\begin{itemize}
    \item First-class labels, with label values $l$ as well as operations for computing on
labels (|canFlowTo|, |lub| and |glb|),
    \item operations for inspecting (\textbf{getLabel}) and modifying (\textbf{setLabel}) the current label of the task (a task can only increase its label, and not decrease it),
    \item operations for non-blocking inter-task communication (\textbf{send} and \textbf{recv}),
        which interact with the global store of per-task message queues |iS|, and
    \item a forking operation allowing to spawn new tasks, which will be described in more detail in the next section.
\end{itemize}

The form of \textbf{send}/\textbf{recv} are \Red{inspired by methods of inter-worker communication
in web browser APIs} \Red{cite}, however in Section~\ref{sec:extensions} we
discuss other possible choices for information flow control aware communication
primitives.  Their rules are different from non-IFC systems
and worth commenting on: when a task places
a message on the message queue of receiving task, it also labels that
message with a label |il'| (which must be above the tasks current label |il|)
When the recipient reads out messages from its queue,
messages whose labels do not
flow to |il| are filtered out (denoted by |filter il Q|).
Specifically, if |Q| is the empty list |nil|, the
function is simply the identity function, i.e.,
|filter il nil = nil|; otherwise:
\[
|filter il ((il', id, ie) , Q )| = \left\{
\begin{array}{l l}
|(il', id, ie) , filter il Q| & \quad \text{if |il' canFlowTo il|}\\
|filter il Q| & \quad \text{otherwise}
\end{array} \right.
\]
This ensures that tasks cannot receive messages that are more secret
than their current label would allow.

@@

\inferrule[I-labelOp]
{ |denot (il1 lop il2) = iv|}
{ |niconf iS id il (il1 lop il2) -> niconf iS id il iv| }
\end{mathpar}
\caption{IFC language with all single-threaded operations}
\label{fig:ifc}
\end{figure}

@@

\Red{TODO[ds]:
  - move the async recv to text and out of the figure
  - clarify discussion on exceptions
  - fix the toLabled
  - add taskIDs to ruls
}
Finally, in Figure~\ref{fig:embedding},
we provide all of the rules responsible for actually carrying out the embedding.
The most important feature of this embedding is that every task maintains its own
copy of the target language global state and evaluation context, thus
enforcing isolation between various tasks.  In more detail:

@@

\begin{itemize}
    \item We extend the values, expressions and evaluation contexts of both languages to allow for embeddings terms in one language to be embedded in the other.
    \item We reinterpret |CE| to be evaluation contexts on task pools, providing definitions for $\mathcal{E}_{\tar{\Sigma}}$ and $\mathcal{E}_{\ifc{\Sigma}}$.  These rules only operate on the first task in the task pool, which by convention, is the only task executing.
    \item We reinterpret |->|, an operation on a single task, in terms of |.->|, operation on task pools.  The correspondence is simple: a task executes a step and then is rescheduled in the task pool according to schedule policy $\alpha$.  Not all schedulers preserve non-interference, so we provide two specific schedules in Figure~\ref{fig:scheduler}.
    \item Finally, we define some rules for scheduling, handling forking tasks (which interact with the state of the target language) and intermediating between the borders of the two languages.
\end{itemize}

The \textsc{I-fork} rule is of particular interest, as it is intended to allow users
to create new tasks.  When a new task is created, there is a question of what the
global state of the new task should be; this is defined in terms of a function $\kappa$.
Conservatively, $\kappa$ may be simply thought of as the identity function, in which
case the semantics of fork are such that the state of the target language is \emph{cloned}
when forking occurs.  However, this is not necessary: it is also valid for $\kappa$
to remove entries from the state.  In Section~\ref{sec:concrete}, we give a more detailed
discussion of the implications of the choice of $\kappa$.

@@

\paragraph{Sending mutable references}  As each task maintains its own
global state for the target language, one might reasonably wonder what would
happen if you attempted to send a reference to another task.

@@

It would be a violation of information flow if such an operation actually
resulted in references being shared between two tasks.  In fact, the address
will be received by the forked task, but it will be a dangling pointer; the
task will become stuck and removed from execution by \textsc{I-noStep}.  (Readers
who are curious whether or not a static discipline could be enforced to
prevent these dangling pointers should check Section~\ref{sec:extensions}.)

@@

When $\kappa$ is the identity, the forked task successfully dereferences
the contents of the reference to be true twice (recall that references
are cloned into the new global state).  When $\kappa$ is the constant function
to the empty store, $x$ becomes a dangling pionter.

@@

At first glance, it does not appear that our system has any such operator;
there are only tasks, and the correct behavior for a task throwing
an exception is self-evident: it should simply be removed from the scheduler.
However, under a \emph{sequential scheduler}, the behavior of \textbf{fork}
can be used to model the behavior of \textbf{toLabeled}:

\begin{align*}
    & |toLabeled l te @= | \\
    & |tab let pid = taskId; | \\
    & |tab let i = TI (fork (IT (TI (setLabel l); TI (send pid l (IT te)))))| \\
    & |tab in (\ _ . TI (setLabel l); TI ( x, _ <= recv . x ))| \\
\end{align*}

When the sequential scheduler is used, the forked task runs to
completion before returning to the task that forked it.  Thus,
a task pool should rather be interpreted as a number of information-flow
aware stack frames.  To read the result of a |toLabeled| operation
(which is communicated via message passing), the parent task must raise
its label; otherwise the label will get dropped, as per the \textsc{I-recv} rule.\footnote{A keen reader may note that with a single, per-task
channel, we might mix up our messages.  Indeed, a proper development requires
the ability to allocate fresh channels; for purposes of simplicity, this
has been omitted.}
